Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data

### 2021-01-11 submission/NatCom_all_models.tex
full pass to revise cv discussion, revise comments, revise response
smooth out a bit

\michael{Refine text in this subsection after we get the final version of all the tables and figures in the appendix.  Things to emphasize: we have a huge amount of data that we are summarizing very succicently; the table here is the tip of the iceberg; the results show our metrics are good, as a sanity check; they show that you can get some results in general, but looking at the details in the appendix show that results are more reliable within a given architecture series than across different architectures; point to out publicly-availab e repo; mention that we do Kendall-$\tau$ rank correlation also.}

We have performed a large-scale analysis of hundreds of publicly-available models.
This broader analysis is on a much larger set of CV and NLP models, with a more diverse set of architectures, that have been developed for a wider range of tasks; and it complements the previous more detailed analysis on CV and NLP models, where we have analyzed only a single architecture series at a time.
See the Supplementary Information
%~\ref{sxn:appendix} 
(and our publicly-available repo)
for details.
To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares (similar results are seen with rank correlation metrics) to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables).
These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
We regress them individually on each of the norm-based and PL-based metrics.

\input{table_3}
\charles{Tabel 3 is autogenerated now: I think R2 adjusted may be a better statistic for us}

Results are summarized in Table~\ref{table:results}.
For the mean, larger $R^{2}$ and smaller $MSE$ are desirable; and for the standard deviation, smaller values are desirable.
Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation;
and PL-based metrics are much better for $MSE$ mean and standard deviation.
These and other results suggest our conclusions hold much more generally.

>>>>

%%\michael{Refine text in this subsection after the appendix is finished.  Things to emphasize: we have a huge amount of data that we are summarizing very succicently; the table here is the tip of the iceberg; the results show our metrics are good, as a sanity check; they show that you can get some results in general, but looking at the details in the appendix show that results are more reliable within a given architecture series than across different architectures; point to out publicly-availab e repo; mention that we do Kendall-$\tau$ rank correlation also.}

We have performed a large-scale analysis of hundreds of publicly-available models.
This broader analysis is on a much larger set of CV and NLP models, with a more diverse set of architectures, that have been developed for a wider range of tasks; and it complements the previous more detailed analysis on CV and NLP models, where we have analyzed only a single architecture series at a time.
See the Supplementary Information
%~\ref{sxn:appendix} 
(and our publicly-available repo)
for details.
To quantify the relationship between quality metrics and the reported test error and/or accuracy metrics, we use ordinary least squares 
%(similar results are seen with rank correlation metrics) 
to regress the metrics on the Top1 (and Top5) reported errors (as dependent variables), and we report the RMSE, the $R^2$ (R2) regresssion metric, and the Kendal-$\tau$ rank correlation metric.
These include Top5 errors for the ImageNet-1K model, percent error for the CIFAR-10/100, SVHN, CUB-200-2011 models, and Pixel accuracy (Pix.Acc.) and Intersection-Over-Union (IOU) for other models.
We regress them individually on each of the norm-based and PL-based metrics.

\input{table_3}

Results are summarized in Table~\ref{table:results} (and Figures~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} of the Supplementary Material).
For the mean, 
smaller RMSE, 
larger $R^2$, and 
larger Kendal-$\tau$ 
are desirable; 
and, for the standard deviation, smaller values are desirable.
Taken as a whole, over the entire corpus of data, PL-based metrics are somewhat better for both the $R^{2}$ mean and standard deviation;
and PL-based metrics are much better for RMSE mean and standard deviation.
Model diagnostics (Supplementary Material) indicate many outliers and imperfect fits.
Overall, though, these and other results suggest our conclusions hold much more generally.

### 2021-01-11 submission/NatCom_appendix.tex
table 5 updated with totals per dataset
smooth out a bit

>%\section{Reproducibility Appendix}
\section{Supplementary Information}
\label{sxn:appendix}

@@

\input{table_6}

@@

\michael{Charles, would you put in a couple of paragraphs about how you generated the tables/figures in the appendix, and a few things we learn from them, i.e., basically enough to convince that reviewer that we paid attention to their concerns.}
\charles{fixed below}
\michael{Table~\ref{table:architectures} still needs to be fixed, since it does not include the overall results partitioned by dataset that we had in the old Table~\ref{table:datasets}. }

@@

Table~\ref{table:results} and Figures~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I} below present results for OLS 
(Ordinary Least Squares) regressions for every dataset and architecture series listed in Table~\ref{table:architectures},
 created by running the \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}
on numerous pretrained models taken from the \texttt{OSMR/imgclsmob Sandbox} github repository of pretrained CV DNN model
ggithub repository. \michael{citation here ?}
In each Figure,  each row of subfigures, for a given pretrained model and dataset, depicts the average Norm-based and Power Law metrics
 ($\langle\log\Vert\cdot\Vert_{F}\rangle$,    $\langle\log\Vert\cdot\Vert_{\infty}\rangle$,
$\hat{\alpha}$,  $\langle\log\Vert\cdot\Vert^{\alpha}_{\alpha}\rangle$)
 against the Top1 Test Accuracy, as reported in the github repository README file~\cite{osmr} ),
along with a shaded area representing the $95\%$ confidence bound.
For each regression, we report the RMSE, R2 regresssion metrics, and Kendal-$\tau$ rank correlation metric,
both in the title of each subfiguire, and in Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
To reproduce these Figures and Tables, see the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook, as listed in Table~\ref{},
and provided github repo accompanying this paper.
The reader may regenerate these Figures and Tables,
and even more fine grained results, by rerunning
 the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook (see Table~\ref{table:notebooks}), which 
analyzes the precomputed data in the \texttt{df\_all.xlsx} file. 
This repository also contains the original \texttt{Google Colab} notebooks, run in January 2020, 
which downloads the pretrained models and  and runs the \texttt{WeightWatcher} tool on them.
%To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 rrors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
%We record the $RMSE$, $R^{2}$, and Kendall-$\tau$ for each metric, averaged over all regressions for all architectures and datasets;
%see Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
%captions fixed in table Ktau.  
%In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
The reader may also run the  \texttt{WeightWatcher} locally on each of the 
pretrained models,
such as the ResNet models, trained on the ImageNet-1K dataset, using the \texttt{WeightWatcher-ResNet-1K.ipynb} notebook.
(However, note that the models may have changed, giving slightly different results).
The final analysis includes 108 regressions in all.%, those with 4 or more models.%, with a positive $R^2$.
See Figure~\ref{fig:summary_regressions_A}-\ref{fig:summary_regressions_I} for more details.
%5Figure~\ref{fig:summary_regressions_B},
%Figure~\ref{fig:summary_regressions_C},
%Figure~\ref{fig:summary_regressions_D},
%Figure~\ref{fig:summary_regressions_E},
%Figure~\ref{fig:summary_regressions_F},
%Figure~\ref{fig:summary_regressions_G},
%Figure~\ref{fig:summary_regressions_H}, and

\michael{Charles, would you put a bit of discussion, e.g., on best and worst, again just a teaser to satisfy that reviewer.}
\charles{below}
From these Figures, we recognize fits of varying quality, ranging from remarkably good to completely uncorrelated.
Starting with the best, consider the Imagenet-1K PreResNet results.  Figure~\ref{fig:summary_regressions_A_10} shows 
the Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$, 
which has a rather large $RMSE=3.93$, and rather small $R2=0.36$, and Kendal-$\tau=0.54$, and
has 6 out of 13 points outside the $95\%$ confidence bands.  In contrast, 
the Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$, 
in  Figure~\ref{fig:summary_regressions_A_12} , 
has a much smaller $RMSE=1.93$, and much larger  $R2=0.85$, and Kendal-$\tau=0.87$, 
and only 2 points outside the $95\%$ confidence bands.
For examples of lower quality fits, consider the SqueezeNext results, as shown Figures~\ref{fig:summary_regressions_C_10} and 
~\ref{fig:summary_regressions_C_12}.  
The Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$ appears
visually anti-correlated with the test accuracies
(as it is with ShuffleNet, in Figure ~\ref{fig:summary_regressions_B_02}).
It has a very large $95\%$ confidence band, with only 2 points close to the regression line, 
a large RMSE,  $R2=0.07$ (i.e zero), and small Kendal-$\tau=0.33$.  
The Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$ is (as always) positively-correlated
with test accuracies, but with $R2=0.43$, show some linear correaltion, and a reasonable Kendal-$\tau=0.73$,
showing moderately strong rank correlation.

>>>>

%\section{Reproducibility Appendix}
\section{Supplementary Information}
\label{sxn:appendix}

@@

\input{table_arch}

@@

@@

Here, we present a more detailed discussion of our large-scale analysis of hundreds of models, which were summarized in Table~\ref{table:results} of Section~\ref{sxn:all_cv_models}. 
We ran the \texttt{WeightWatcher} tool (version 0.2.7)~\cite{weightwatcher_package}
on numerous pretrained models taken from the \texttt{OSMR/imgclsmob Sandbox} github repository of pretrained CV DNN models~\cite{osmr},   
performing OLS (Ordinary Least Squares) regressions for every dataset and architecture series listed in Table~\ref{table:architectures}.
Table~\ref{table:results} summarized the overall results, and Figures~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} below present a more detailed visual summary, which enables model diagnostics.

For each Figure, each row of subfigures considers a given pretrained model and dataset, depicting the average Norm-based and Power Law 
metrics---Log Frobenius norm
          ($\langle\log\Vert\mathbf{W}\Vert^{2}_{F}\rangle$),
          Log Spectral norm
          ($\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$),
          Weighted Alpha
          ($\hat{\alpha}$),
          and Log $\alpha$-Norm
          ($\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$)---against 
the Top1 Test Accuracy, as reported in the github repository README file~\cite{osmr}, along with a shaded area representing the $95\%$ confidence bound.
For each regression, we report the RMSE, the R2 regresssion metric, and the Kendal-$\tau$ rank correlation metric in the title of each subfigure.
We also present these same numerical values in Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
To reproduce these Figures and Tables, see the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook, as listed in Table~\ref{table:notebooks}.
(These are provided in the github repo accompanying this paper.)
The reader may regenerate these Figures and Tables, as well as more fine grained results, by rerunning the \texttt{OSMR-Analysis.ipynb} python Jupyter Notebook (see Table~\ref{table:notebooks}), which analyzes the precomputed data in the \texttt{df\_all.xlsx} file. 
This repository also contains the original \texttt{Google Colab} notebooks, run in January 2020, which download the pretrained models and run the \texttt{WeightWatcher} tool on them.
%To generate the results in Table~\ref{table:results}, we run linear regressions for each architecture series in Table~\ref{table:architectures}, regressing each empirical Log Norm metric against the reported Top1 rrors (as listed on the \texttt{osmr/imgclsmob} github repository README file~\cite{osmr}, with the relevant data extracted and provided in our github repo as \texttt{pytorchcv.html}).
%We record the $RMSE$, $R^{2}$, and Kendall-$\tau$ for each metric, averaged over all regressions for all architectures and datasets;
%see Table~\ref{table:RMSEresults}, Table~\ref{table:R2results}, and Table~\ref{table:Ktauresults}, respectively.
%captions fixed in table Ktau.  
%In the repo, plots are provided for every regression, and more fine grained results may be computed by the reader by analyzing the data in the \texttt{df\_all.xlsx} file.
The reader may also run the  \texttt{WeightWatcher} locally on each of the 
pretrained models,
such as the ResNet models, trained on the ImageNet-1K dataset, using the \texttt{WeightWatcher-ResNet-1K.ipynb} notebook.
(We should note, however, that the publicly-available versions of these models may have changed slightly, giving slightly different results).
Our final analysis includes 108 regressions in all. %, those with 4 or more models.%, with a positive $R^2$.
See Figure~\ref{fig:summary_regressions_A}--\ref{fig:summary_regressions_I} for more details.
%5Figure~\ref{fig:summary_regressions_B},
%Figure~\ref{fig:summary_regressions_C},
%Figure~\ref{fig:summary_regressions_D},
%Figure~\ref{fig:summary_regressions_E},
%Figure~\ref{fig:summary_regressions_F},
%Figure~\ref{fig:summary_regressions_G},
%Figure~\ref{fig:summary_regressions_H}, and

% Log Frobenius norm
% Log Spectral norm
% Weighted Alpha
% Log $\alpha$-Norm

From these Figures, we recognize fits of varying quality, ranging from remarkably good to completely uncorrelated.
Starting with some of the best, consider the Imagenet-1K PreResNet results.  
For example,
Figure~\ref{fig:summary_regressions_A_10} shows the 
Log Spectral norm, %Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$, 
which has a rather large $RMSE=3.93$, and rather small $R2=0.36$, and Kendal-$\tau=0.54$, and 
which has 6 out of 13 points outside the $95\%$ confidence bands.  
In contrast, the 
Log Frobenius norm %Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$, 
in Figure~\ref{fig:summary_regressions_A_12}
has a much smaller $RMSE=1.93$, a much larger $R2=0.85$, and Kendal-$\tau=0.87$, and 
has only 2 points outside the $95\%$ confidence bands.
For examples of lower quality fits, consider the SqueezeNext results, as shown Figures~\ref{fig:summary_regressions_C_10} and~\ref{fig:summary_regressions_C_12}.  
The 
Log Spectral norm  %Avg. Log Spectral Norm $\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$ 
appears visually anti-correlated with the test accuracies (as it is with ShuffleNet, in Figure~\ref{fig:summary_regressions_B_02}).
It has a very large $95\%$ confidence band, with only 2 points close to the regression line, a large RMSE, $R2=0.07$ (i.e., near zero), and small Kendal-$\tau=0.33$.  
The 
Log Frobenius norm %Avg. Log Alpha Norm $\langle\log\Vert\mathbf{W}\Vert^{\alpha}_{\alpha}\rangle$ 
is (as always) positively-correlated with test accuracies, but with $R2=0.43$, it show some linear correlation, and a reasonable Kendal-$\tau=0.73$, showing moderately strong rank correlation.

Many more such conclusions can be drawn by examining these Tables and Figures and reproducing the results from our publicly-available repo.

### 2021-01-11 submission/NatCom_cv_models.tex
smooth out a bit

We examine the performance of the four quality metrics (Log Frobenius norm, Log Spectral norm, Weighted Alpha, and Log $\alpha$-Norm) applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
%
Figure~\ref{fig:vgg-metrics} plots the four quality metrics versus reported test accuracies~\cite{pytorch},% 
\footnote{These test accuracies have been previously reported and made publicly-available by others.  We take them as given.  We do not attempt to reproduce/verify them, since we do not permit ourselves access to training/test data.}
as well as a basic linear regression line, for the VGG series.
Here, smaller norms and smaller values of $\hat{\alpha}$ imply better generalization (i.e., greater accuracy, lower error). 
Quantitatively, Log Spectral norm is the best; but, visually, all four metrics correlate quite well with reported Top1 accuracies.
The DenseNet series has similar behavior.
(These and many other such plots can be seen on our publicly-available~repo.)

@@

\michael{Charles, any more sentences to add to the last three paragraphs to clarify anything?  If not, then just comment out this comment.}
\charles{In partcular, the PL-based AlphaHat $\hat{\alpha}$ and Shatten-Norm $\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$ metrics 
tend to perform better when there is a wider variation in the hyper-parameters going beyond just increasing the depth.  In fact sometimes the
purely norm-based metrics such as the Spectral Norm can be uncorrelated or even
anti-correlated with the test accuracy, as seen in Section~\ref{} with
ShuffleNet in Figure~\ref{fig:summary_regressions_B_02},
SqueezeNext in Figure~\ref{fig:summary_regressions_C_10},
and WRN in Figure~\ref{fig:summary_regressions_G_06}).
}

>>>>

We examine the performance of the four quality metrics---Log Frobenius norm
                                                         ($\langle\log\Vert\mathbf{W}\Vert^{2}_{F}\rangle$), 
                                                         Log Spectral norm
                                                         ($\langle\log\Vert\mathbf{W}\Vert^{2}_{\infty}\rangle$),
                                                         Weighted Alpha
                                                         ($\hat{\alpha}$),
                                                         and Log $\alpha$-Norm
                                                         ($\langle\log\Vert\mathbf{X}\Vert^{\alpha}_{\alpha}\rangle$)---applied to each of the VGG, ResNet, ResNet-1K, and DenseNet series.
%
Figure~\ref{fig:vgg-metrics} plots the four quality metrics versus reported test accuracies~\cite{pytorch},% 
\footnote{These test accuracies have been previously reported and made publicly-available by others.  We take them as given.  We do not attempt to reproduce/verify them, since we do not permit ourselves access to training/test data.}
as well as a basic linear regression line, for the VGG series.
Here, smaller norms and smaller values of $\hat{\alpha}$ imply better generalization (i.e., greater accuracy, lower error). 
Quantitatively, Log Spectral norm is the best; but, visually, all four metrics correlate quite well with reported Top1 accuracies.
The DenseNet series has similar behavior.
(These and many other such plots can be seen on our publicly-available~repo.)

@@

In particular, the PL-based 
Weighted Alpha and Log $\alpha$-Norm
metrics tend to perform better when there is a wider variation in the hyperparameters, going beyond just increasing the depth.  
In addition, sometimes the purely norm-based metrics such as the Log Spectral norm can be uncorrelated or even anti-correlated with the test accuracy, while the PL-metrrics are positively-correlated.
This is seen in the Supplementary Material, 
ShuffleNet in Figure~\ref{fig:summary_regressions_B_02},
SqueezeNext in Figure~\ref{fig:summary_regressions_C_10}, and 
WRN in Figure~\ref{fig:summary_regressions_G_06}).
